Bioinformatics analysis of whole-genome bisulfite-sequencing data

### 2014-12-03 Rmarkdown/wgbs_bioinformatics_analysis.Rmd
Keep on truckin
Misc. notes.
Fixed some syntax errors

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Bis-SNP.pdf}
\caption{`Bis-SNP` is able to distinguish unmethylated cytosines (site 1), from cytosine to thymine genetic variants (site 2) and thymine to (unmethylated) cytosine genetic variants (site 3) by examining the reads mapped to the reverse strand. For all three loci, the reads mapped to the forward strand contain a thymine. However, it is the base on the reverse strand that reveal the true genotype. When combined with the reference genome it can be inferred whether the sample's genome, which isn't directly observed, has a genetic variant at that location. This is only possible with bisulfite-data generated using the directional protocol. This figure is adapted from \cite{Liu:2012ge}}
\label{fig:Bis-SNP.pdf}
\end{figure}

@@

@@

\begin{table}[h]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
Software           & Reference                                & Input                                              & Designed for targeted sequencing & Other limitations                                                                \\ \midrule
\code\{methclone\} & \cite\{Li:2014ei\}                       & Bismark BAM files                                  & No, although example uses eRRBS  & \textbf\{TODO: Try installing. Suspect you can't get actual methylation calls.\} \\
\code\{methpat\}   & \url\{https://github.com/bjpop/methpat\} & Output of \code\{bismark\_methylation\_extractor\} & Yes (amplicons)                  &                                                                                  \\
\code\{DMEAS\}     & \cite\{He:2013cj\}                       & Output of \code\{bismark\_methylation\_extractor\} &                                  & Windows operating system only. Perl code only available as PDF file.             \\ \bottomrule
\end{tabular}
\caption{My caption}
\label{tab:other_m-tuple_methylation_callers}
\end{table}

@@

```{r methtuple_runtime, eval = TRUE, echo = FALSE, message = FALSE, fig.lp = "fig:methtuple_runtime", fig.cap = "The running times are the 'User time' reported by \code{GNU time} converted from seconds to minutes. The suffix 'ac' on the tuple size means that the option \texttt{--all-combinations} was set.", fig.width = 8, dev = "pdf"}
library(dplyr)
library(ggplot2)
library(stringr)
library(gridExtra)

@@

```{r methtuple_memory, eval = TRUE, echo = FALSE, message = FALSE, fig.lp = "fig:methtuple_memory", fig.cap = "; the suffix 'ac' means that the option \texttt{--all-combinations} was set. This option creates all tuples with $NIL \geq 0$ which is far greater than the equivalent tuples with $NIL = 0$. The maximum memory usage is the 'Maximum resident set size' reported by \texttt{GNU time} converted from kilobytes to gigabytes. These values are divided by four to fix bug in how \texttt{GNU time} reports the maximum memory usage (\url{https://bugzilla.redhat.com/show_bug.cgi?id=703865})", fig.width = 8, dev = "pdf"}
# Memory
memory <- lapply(paste0("../data/methtuple_performance/memory/", 
                      list.files("../data/methtuple_performance/memory/")), 
               read.table)
names(memory) <- gsub(".memory", "", 
                    list.files("../data/methtuple_performance/memory/"))
n <- str_split(names(memory), "\\.")
# NB: Need to divide reported memory usage by 4 due to bug in 'time'
# (https://bugzilla.redhat.com/show_bug.cgi?id=703865)
memory <- data_frame("Chromosome" = factor(rep(sapply(n, "[", 1), 
                                             times = sapply(memory, nrow)),
                                         levels = 
                                           paste0("chr", 
                                                  c(1:22, "X", "Y", "M", "L"))),
                   "Tuple" = rep(sapply(n, "[", 2),
                                      times = sapply(memory, nrow)),
                   "Memory" = unlist(memory) / 4)
gg_memory <- memory %>%
  ggplot(aes(x = Chromosome, y = Memory / 10^6)) + geom_boxplot() +
  facet_wrap( ~ Tuple, ncol = 3) + thesis_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ylab("Memory usage (GB)") + 
  ggtitle("Maximum memory usage of methtuple across 49 samples")
facet_wrap_labeller(gg_memory, paste0("Tuple size: ", unique(memory$Tuple)))
```

>>>>

\begin{figure}[h]
\includegraphics[width=\textwidth]{../figures/Bis-SNP.pdf}
\caption{`Bis-SNP` is able to distinguish unmethylated cytosines (site 1), from cytosine to thymine genetic variants (site 2) and thymine to (unmethylated) cytosine genetic variants (site 3) by examining the reads mapped to the reverse strand. For all three loci, the reads mapped to the forward strand contain a thymine. However, it is the base on the reverse strand that reveal the true genotype. When combined with the reference genome it can be inferred whether the sample's genome, which isn't directly observed, has a genetic variant at that location. This is only possible with bisulfite-data generated using the directional protocol. This figure is adapted from \cite{Liu:2012ge}.}
\label{fig:Bis-SNP}
\end{figure}

@@

__TODO: Size of m-tuples histogram__

@@

\begin{table}[h]
\centering
\begin{tabular}{@{}lllll@{}}
\toprule
Software           & Reference                                & Input                                              & Designed for targeted sequencing & Other limitations                                                                \\ \midrule
\texttt\{methclone\} & \cite\{Li:2014ei\}                       & Bismark BAM files                                  & No, although example uses eRRBS  & \textbf\{TODO: Try installing. Suspect you can't get actual methylation calls.\} \\
\texttt\{methpat\}   & \url\{https://github.com/bjpop/methpat\} & Output of \texttt\{bismark\_methylation\_extractor\} & Yes (amplicons)                  &                                                                                  \\
\texttt\{DMEAS\}     & \cite\{He:2013cj\}                       & Output of \texttt\{bismark\_methylation\_extractor\} &                                  & Windows operating system only. Perl code only available as PDF file.             \\ \bottomrule
\end{tabular}
\caption{My caption}
\label{tab:other_m-tuple_methylation_callers}
\end{table}

@@

```{r methtuple_runtime, eval = TRUE, echo = FALSE, message = FALSE, fig.lp = "fig:methtuple_runtime", fig.cap = "The running times are the 'User time' reported by \texttt{GNU time} converted from seconds to minutes. The suffix 'ac' on the tuple size means that the option \texttt{--all-combinations} was set.", fig.width = 8, dev = "pdf"}
library(dplyr)
library(ggplot2)
library(stringr)
library(gridExtra)

@@

```{r methtuple_memory, eval = TRUE, echo = FALSE, message = FALSE, fig.lp = "fig:methtuple_memory", fig.cap = "The maximum memory usage is the 'Maximum resident set size' reported by \texttt{GNU time} converted from kilobytes to gigabytes. These values are divided by four to fix bug in how \texttt{GNU time} reports the maximum memory usage (\\url{https://bugzilla.redhat.com/show_bug.cgi?id=703865}). The suffix 'ac' means that the option \texttt{--all-combinations} was set.", fig.width = 8, dev = "pdf"}
# Memory
memory <- lapply(paste0("../data/methtuple_performance/memory/",
                      list.files("../data/methtuple_performance/memory/")),
               read.table)
names(memory) <- gsub(".memory", "",
                    list.files("../data/methtuple_performance/memory/"))
n <- str_split(names(memory), "\\.")
# NB: Need to divide reported memory usage by 4 due to bug in 'time'
# (https://bugzilla.redhat.com/show_bug.cgi?id=703865)
memory <- data_frame("Chromosome" = factor(rep(sapply(n, "[", 1),
                                             times = sapply(memory, nrow)),
                                         levels =
                                           paste0("chr",
                                                  c(1:22, "X", "Y", "M", "L"))),
                   "Tuple" = rep(sapply(n, "[", 2),
                                      times = sapply(memory, nrow)),
                   "Memory" = unlist(memory) / 4)
gg_memory <- memory %>%
  ggplot(aes(x = Chromosome, y = Memory / 10^6)) + geom_boxplot() +
  facet_wrap( ~ Tuple, ncol = 3) + thesis_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + ylab("Memory usage (GB)") +
  ggtitle("Maximum memory usage of methtuple across 49 samples")
facet_wrap_labeller(gg_memory, paste0("Tuple size: ", unique(memory$Tuple)))
```
